Spot: a tool to locate numerical differences in pipelines

### 2019-12-16 paper.tex

\title{NURM: a tool to 
locate numerical differences in pipelines}

@@

There are different approaches to improve the reproducibility of the analysis, 
including containerization techniques that encapsulate software/hardware dependencies, 
provenance capturing tools, and version control systems. 
However, a comprehensive solution requires to fix the numerical instabilities instead of 
masking the problem. For this purpose, bagging is one solution that has been used to 
stabilize motion estimation analysis in fMRI~\cite{Glatard2018hbm}. 
Bootstrap aggregation reduces the instabilities, but it needs expensive computations.
As the second solution, a debugging tool can help to identify and 
fix the instabilities in the pipeline. Therefore, we introduce 
the Numerical Reproducibility Measurement (NURM) tool to identify 
the processes in the pipeline that create numerical differences across different conditions. 
Furthermore, we can stabilize the identified processes in the next steps.

The remainder of this paper describes the NURM-tool in different parts 
including the way of capturing and representation of the provenance 
information, the method of clustering of different subject types, and 
process labelling to characterize differences. We describe the HCP 
pre-processing pipelines as a popular neuroimaging project to test the 
NURM-tool and evaluate its functionality. Finally, we reported the 
experimental results which describe the processes that are responsible 
for the differences in the pipelines.

@@

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=1\columnwidth]{images/fig1-a}
    \caption{Identification of temporary and multi-write files.
    (\href{https://docs.google.com/drawings/d/1rsbFwPjNPNMRhXUl4RSK3G0OaiJC7K-buiHN0hPD8eQ/edit?usp=sharing}{Source})}
    \label{fig:overview-capturing}
  \end{subfigure}
   \begin{subfigure}{\columnwidth}
    \centering
     \includegraphics[width=1\columnwidth]{images/fig1-b}
     \caption{Labelling process graphs including intermediary files.
     (\href{https://docs.google.com/drawings/d/1yblYuKWAD18aJe5JxBu2h1EysyoQA3YrsQj-1T3s0l8/edit?usp=sharing}{Source})}
     \label{fig:overview-labelling}
   \end{subfigure}
   \caption{An overview of NURM-tool in two steps. First capturing intermediary files (a); then
            labelling process graphs (b).}
   \label{fig:overview}
  \end{figure}

@@

In particular, NURM-tool is tested on the neuroimaging applications which are 
predominantly using mathematical libraries. Therefore, we expect to find 
differences as a result of changing the mathematical functions between operating system libraries.

@@

NURM-tool identified \emph{mris\_make\_surfaces} as the only process in 
the pipeline that creates differences in 10 subjects among 20 subjects.
This process generates surface files for cortical and white matter surfaces 
(see~\cite{fischl2012freesurfer} for more detail). 
Also, FreeSurfer did not introduce any differences in 10 subjects.

>>>>

\title{Spot: a tool to 
locate numerical differences in pipelines}

@@

There are different approaches to improve the reproducibility of the analysis, 
including containerization techniques that encapsulate software/hardware dependencies, 
provenance capturing tools, and version control systems. 
However, a comprehensive solution requires to fix the numerical instabilities instead of 
masking the problem. For this purpose, bagging is one solution that has been used to 
stabilize motion estimation analysis in fMRI~\cite{Glatard2018hbm}. 
Bootstrap aggregation reduces the instabilities, but it needs expensive computations.
As the second solution, a debugging tool can help to identify and 
fix the instabilities in the pipeline. Therefore, we introduce 
the numerical reproducibility measurement tool named \emph{Spot} to identify 
the processes in the pipeline that create numerical differences across different conditions. 
Furthermore, we can stabilize the identified processes in the next steps.

The remainder of this paper describes the Spot tool in different parts 
including the way of capturing and representation of the provenance 
information, the method of clustering of different subject types, and 
process labelling to characterize differences. We describe the HCP 
pre-processing pipelines as a popular neuroimaging project to test the 
Spot tool and evaluate its functionality. Finally, we reported the 
experimental results which describe the processes that are responsible 
for the differences in the pipelines.

@@

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=1\columnwidth]{images/fig1-a}
    \caption{Identification of temporary and multi-write files.
    (\href{https://docs.google.com/drawings/d/1rsbFwPjNPNMRhXUl4RSK3G0OaiJC7K-buiHN0hPD8eQ/edit?usp=sharing}{Source})}
    \label{fig:overview-capturing}
  \end{subfigure}
   \begin{subfigure}{\columnwidth}
    \centering
     \includegraphics[width=1\columnwidth]{images/fig1-b}
     \caption{Labelling process graphs including intermediary files.
     (\href{https://docs.google.com/drawings/d/1yblYuKWAD18aJe5JxBu2h1EysyoQA3YrsQj-1T3s0l8/edit?usp=sharing}{Source})}
     \label{fig:overview-labelling}
   \end{subfigure}
   \caption{An overview of Spot tool in two steps. First capturing intermediary files (a); then
            labelling process graphs (b).}
   \label{fig:overview}
  \end{figure}

@@

In particular, Spot tool is tested on the neuroimaging applications which are 
predominantly using mathematical libraries. Therefore, we expect to find 
differences as a result of changing the mathematical functions between operating system libraries.

@@

Spot tool identified \emph{mris\_make\_surfaces} as the only process in 
the pipeline that creates differences in 10 subjects among 20 subjects.
This process generates surface files for cortical and white matter surfaces 
(see~\cite{fischl2012freesurfer} for more detail). 
Also, FreeSurfer did not introduce any differences in 10 subjects.
